\section{Conclusions}
In this paper, we present a general algorithmic framework for lifting 
the model search and conflict analysis procedures in satisfiability 
solvers to program analysis.  We embody these techniques in a tool, ACDLP,  
for automatic bounded safety verification of C programs over a template 
polyhedra abstract domains.  

We present an {\em abstract model search} procedure that uses a 
parameterised abstract transformer to flexibly control the precision and 
efficiency of the deductions in the template polyhedra abstract domain. 
The underlying expressivity of the abstract domain helps our decision 
heuristics to exploit the high-level structure of the program for making 
effective decisions.  The {\em abstract conflict analysis} procedure learns 
abstract transformers over a given template following a UIP computation.
Experimental evaluation over a range of benchmarks shows~10X
reduction in the total number of {\em decisions}, {\em propagations}, 
{\em conflicts} and {\em backtracking} iterations compared to CBMC.  Moreover, 
ACDLP is \rmcmt{?X} faster than CBMC.  Compared to Astr{\'e}e, ACDLP is
\rmcmt{70\%} more precise on our benchmark suite.  \rmcmt{In the future, we plan to
extend our framework to unbounded verification through invariant generation.} 


%able to exploit the expressivity of richer abstract domains 
%as well as the precision of the CDCL architecture.  \tmcmt{I don't understand this ``as well as'' business. Do we mean ``without giving up the precision of...''?}
%\Omit{
%The ACDLP analyses in this paper perform bounded model checking. Since we use abstract 
%domains, however, we will be able to extend it to unbounded verification, a possibility 
%we intend to explore in future work.
%}
