\paragraph{Contributions}
In this paper, we make the following technical contributions.
%
\begin{compactenum}
\item We present a novel program analysis technique called {\em Abstract 
Conflict Driven Learning for Programs} (ACDLP) that lifts the CDCL algorithm 
over template polyhedra abstract domains for safety verification of C programs.  
%We have implemented our analyzer ACDLP for safety verification of C programs. 
We compare the performance of ACDLP compared to a precise model checker CBMC and a 
commercial abstract interpreter Ast{\'r}ee.  Our tool is available 
online\footnote{http://www.cprover.org/acdcl/}. 

\Omit{  
We present a generic implementation framework for model search, conflict analysis, 
learning and backjumping phase that can be instantiated over any arbitrary 
abstract domains, thus making our tool easily extendable to domain specific decision, 
propagation and learning heuristics.  
}
\item We present an abstract model search algorithm that strictly 
generalises the Boolean Constraint Propagation (BCP) procedure. We 
propose several domain specific propagation and decision heuristics 
that infers abstract domain elements over the template polyhedra domain.   
\Omit{
We propose a domain-specific decision heuristics that exploit the 
expressivity of the underlying abstract domain to make effective decisions. 
}
\item We sketch an abstract conflict analysis procedure that operates on an 
abstract conflict graph to learn a generalised reason for conflicts.  
 
\item  Experiments indicate a two order of magnitude reduction 
in the number of decisions, propagations, and conflicts compared 
to a SAT solver while being more precise than an abstract interpreter.  
\Omit {
set of benchmarks drawn from the bitvector regression category in 
SV-COMP, bit-precise software models auto-generated from hardware 
circuits domain and some hand crafted example show that 
}
\end{compactenum}
