\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
%
Static program analysis with abstract interpretation~\cite{CC77} is widely
used to verify properties of safety-critical systems.  Static analyses
commonly aim to compute program invariants as fixed-points of abstract
transformers.  Abstract states are chosen from a lattice that has meet
$(\sqcap)$ and join $(\sqcup)$ operations; the meet precisely models set
intersection (or conjunction, taking a logical view), and the join
over-approximates set union (or disjunction).  Over-approximation in the
join operation is one of the sources of precision loss, which causes false
alarms.  Typical abstract domains are non-distributive; suppose $a$ and $b$
together represent the abstract semantics of a program and $c$ represents a
set of abstract behaviours that violate the specification.  In a
non-distributive domain, $(a \sqcup b) \sqcap c$ can be strictly less
precise than $(a \sqcap c) \sqcup (b \sqcap c)$.  This means that in typical
abstract domains, analysing program behaviours separately can improve the
precision of the analysis.  Usual means to address false alarms therefore
include not only the use of richer abstract domains, but also of refinements
that delay joins or perform some form of case-splitting.  Such techniques
trade off higher precision against lower efficiency and may be susceptible
to case enumeration behaviour.

%%%%%%%%%%%%%%%%%%%%%%% CDCL %%%%%%%%%%%%%%%%%%%%%
By contrast, Model Checking (MC)~\cite{mc-book} can be seen to operate on
distributive lattice structures that represent disjunction without loss
of precision.  Classical MC directly operates on distributive
representations, such as BDDs, while more recent implementations use SAT
solvers.  SAT solvers themselves operate on partial assignments, which are
non-distributive structures.  To handle disjunction, case-splitting is
performed~\cite{sas12}.  Propositional SAT solvers solve large formulae, and
are often able to avoid enumerating cases.  The impressive performance of
modern solvers is credited to well-tuned decision heuristics and
sophisticated clause learning algorithms.  Collectively, these algorithms
are referred to as \emph{Conflict Driven Clause Learning}
(CDCL)~\cite{cdcl}.  An obvious idea is to lift CDCL from the domain of 
partial assignments to other non-distributive domains.

%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
Abstract Conflict Driven Clause Learning (ACDCL)~\cite{dhk2013-popl} is one
such lattice-based generalization of CDCL.  ACDCL is a general algorithmic
framework, parameterized by a concrete domain~$C$ and an abstract
domain~$A$.  Classical CDCL can be viewed as an instance of ACDCL in which
$C$ is the set of propositional truth assignments and $A$ the domain of
propositional partial assignments~\cite{leo-thesis}.  Since the concrete
domain of interest is a parameter to the framework, ACDCL can in principle
be used to build both \emph{logical decision
procedures}~\cite{DBLP:journals/fmsd/BrainDGHK14} and \emph{program
analyzers}.  In the former case, the concrete domain is the set of candidate
models for the formula; in the latter case, it is the set of program traces
that may lead to an error.  Haller
et~al.~in~\cite{DBLP:journals/fmsd/BrainDGHK14} illustrate the first idea by
presenting a floating-point decision procedure that uses interval constraint
propagation.

%%%%%%%%%%%%%%%%%%%%%%% ACDLP %%%%%%%%%%%%%%%%%%%%%

In this paper, we explore the second idea by presenting an extension of 
ACDCL to program analysis.  We call our 
framework \emph{Abstract Conflict Driven Learning for Programs}
(ACDLP).  The key insight of ACDLP is to use decisions and learning to precisely 
reason about disjunctions in non-distributive domains, thereby automatically 
refining the precision of analysis for safety checking of C programs.  
We~introduce two central components of our framework: an abstract
model search algorithm that uses decisions and propagations to 
search for counterexample trace and an abstract conflict analysis 
procedure that approximates a set of unsafe traces through transformer 
learning.
%%%%%%%%%%%%%%%%%%%%%%% Domain %%%%%%%%%%%%%%%%%%%%%
We illustrate the application of our framework to program analysis 
using a \textit{template polyhedra abstract domain}~\cite{vmcai05}, 
which includes most of the commonly used abstract domains, such as boxes, 
octagons, zones and TCMs.  

We present an experimental evaluation of our analyser compared 
to CBMC~\cite{cbmc.tacas:2004}, which uses propositional solvers, and to 
Astr{\'e}e~\cite{DBLP:conf/pldi/BlanchetCCFMMMR03}, a commercial abstract 
interpretation tool.  In this paper, we make the following contributions.

\Omit{
Our experiments suggest that ACDLP can be seen as a 
technique to improve efficiency of SAT-based BMC. Additionally, it can 
also be perceived as an automatic way to improve the precision of abstract 
interpretation for bounded unwindings of programs.  
}
%
\begin{enumerate}
\item A novel program analysis framework that lifts model search and conflict
analysis procedures of CDCL algorithm over a template polyhedra abstract domain. These techniques
are embodied in our tool, \emph{ACDLP}, for automatic bounded safety verification 
    of C programs.
\item A parameterized abstract transformer that guides the model
  search in forward, backward and multi-way direction for counterexample
    detection. 
\item A conflict analysis procedure that performs UIP-based transformer learning 
  over template polyhedra abstract domain through abductive reasoning. 
\end{enumerate}
%

\Omit{
While the focus of this paper is the analysis of loop-free unwindings of programs, our technique is in 
principle compatible with abstract-interpretation-based handling of loops using 
widenings~\cite{leo-thesis}.  Although, finding non-trivial trace abstraction 
that satisfy properties of ACDCL is an open problem.  Finally, we present the 
results of experimental evaluation of our analyser compared to CBMC~\cite{cbmc.tacas:2004}, 
which uses propositional solvers, and to Astr{\'e}e~\cite{DBLP:conf/pldi/BlanchetCCFMMMR03}, 
a commercial abstract interpretation tool. 
}
%%%%%%%%%%%%%%%%%%%%%%% Experiments %%%%%%%%%%%%%%%%%%%%%
%\vspace{-7mm}

% This abstract domain captures many
% relationships between program variables that intervals cannot represent.
% TM - INTERVALS don't express relationships at all. 

%Fundamentally, the abstract domain of polyhedra~\cite{polyhedra} is a 
%complete relational domain that capture all linear inequalities 
%between program variables.  
%It is a `relational' domain, in the sense that
%its elements express relationships between two or more program variables. 
%By contrast, the interval domain expresses constraints on individual program variables.


\Omit {


%This enables an ACDCL-based analyser to automatically refine an imprecise 
%analysis and thus prevents enumeration behaviour.  

%To date, ACDCL has been instantiated as a decision procedure for the
%first-order theory of floating-point arithmetic~\cite{DBLP:journals/fmsd/BrainDGHK14}, 

%\rmcmt{state other limitations}.  
that is, given abstract domain elements $a$, $b$, $c$, 
$(a \sqcup b) \sqcap c \sqsupseteq (a \sqcap c) \sqcup (b \sqcap c)$.

We provide a theoretical recipe and a practical instantiation for 
generalising CDCL architecture to arbitrary abstract domains.   

\paragraph{Abstract Domain.} 
Limited expressivity of abstract domains may lead to imprecise 
over-approximation which gives rise to false alarms. Evidently, the 
verification of many programs requires more expressive numerical domains, 
such as polyhedra for analysis.  These domains capture numerical 
relationship between program variables that intervals cannot express.  
To this end, we provide a theoretical recipe and a practical instantiation 
for generalising CDCL architecture to arbitrary abstract domains, thus 
lifting it to richer lattice structures. 

The key motivation of our work is to combine the precision of a SAT solver 
and the efficiency of an abstract interpreter to present a new class of 
program analysers.  

\rmcmt{Introduce ACDLP here with a distinct notion for programs}
Whereas the propositional reasoning of SAT solvers can be lifted easily from the 
Boolean lattice to other non-relational domains, application to relational 
domains is challenging because of the relational properties of inferred 
deductions and the complexity of the closure operation in fixed-point computation 
during the propagation phase.  In this paper, we identify specific modifications 
to the CDCL algorithm that are necessary to lift propositional CDCL to 
template polyhedra domain. 

\rmcmt{Present program and property driven trace partitioning (with inline example with octagons)} 
}
