% 16 pages in LNCS format including references. Appendices separate. 
\documentclass[a4paper]{llncs}


\linespread{0.95}
\usepackage{alltt}
\usepackage{import}
\usepackage{times}
\usepackage{multirow}
\usepackage{microtype}
\usepackage{array}
\usepackage{graphicx,wrapfig}
\usepackage[noadjust]{cite}
\usepackage{tikz}
\usetikzlibrary{plotmarks}
\usepackage{pgfplotstable}
\usepackage{filecontents}
\usepackage{pgfplots}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{psfrag}
\usepackage{epsfig}
\usepackage{wasysym} 
%\usepackage{subfigure}
\usepackage{paralist}
%\usepackage{dingbat}
\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
\usepackage{hyperref}
%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
\usepackage[subnum]{cases}
\makeatother

\newcommand{\extendedonly}[1]{}
\newcommand{\paperonly}[1]{#1}
\newtheorem{exmp}{Example}

\newcommand{\tool}[1]{\textsc{#1}\xspace}
\newcommand{\cbmcv}{\tool{cbmc 5.0}}
\newcommand{\Omit}[1]{}
\newcommand{\mydef}[1]{\begin{definition}#1\end{definition}}
\newcommand{\rmcmt}[1]{{\color{magenta}{#1}}}%#1
\newcommand{\pscmt}[1]{{\color{blue}{#1}}}%#1
\newcommand{\dkcmt}[1]{{\color{red!70!black}{#1}}}%#1
\newcommand{\tmcmt}[1]{{\color{orange}{#1}}}%#1
\newcommand{\lhcmt}[1]{{\color{purple}{#1}}}%#1

\input{symbols}

\begin{document}

\title{Lifting CDCL to Template-based Abstract Domains\\ for Program Verification}

\author{Rajdeep Mukherjee\inst{1} \and Peter Schrammel\inst{2} \and 
Leopold Haller\inst{3} \and \\ 
Daniel Kroening\inst{1} \and Tom Melham\inst{1}}

\institute{University of Oxford, UK \and University of Sussex, UK \and Google Inc., USA}

\maketitle

%===============================================================================
\begin{abstract}
%===============================================================================
%
The success of Conflict Driven Clause Learning (CDCL) for Boolean
satisfiability has inspired its adoption in other domains.  We present a
novel lifting of the CDCL to program analysis called \emph{Abstract Conflict
Driven Learning for Programs} (ACDLP).  ACDLP alternates between \emph{model
search}, which performs over-approximate deduction with constraint
propagation, and \emph{conflict analysis}, which performs under-approximate
abduction with heuristic choice.  We instantiate the model search and
conflict analysis algorithms to an abstract domain of \textit{template
polyhedra}, strictly generalising CDCL from the Boolean lattice to richer
lattice structures.  Our domain of template polyhedra can express intervals,
octagons, and certain restricted polyhedral constraints over program
variables.  We have implemented ACDLP for automatic bounded safety verification 
of C programs.  We~evaluate the performance of our
analyser by comparing with the CBMC program verifier, which uses
propositional solvers, and Astr{\'e}e, a commercial abstract
interpretation tool.  We observe one order of magnitude reduction in the
number of decisions, propagations, and conflicts as well as~1.5X speedup 
in runtimes for ACDLP compared to CBMC.  Compared to Astr{\'e}e, ACDLP is {63}\% 
more precise.  To the best of our knowledge, this is the first instantiation 
of CDCL with a template polyhedra abstract domain for safety verification.
%
\end{abstract}

%===============================================================================
\input{introduction}
%===============================================================================

%===============================================================================
%\input{contributions}
%===============================================================================

%===============================================================================
\input{example}
%===============================================================================

%===============================================================================
\input{domains}
%===============================================================================

%===============================================================================
\input{algorithm}
%===============================================================================

%===============================================================================
\input{deductions}
%===============================================================================

%===============================================================================
\input{decision}
%===============================================================================

%===============================================================================
\input{learning}
%===============================================================================

%===============================================================================
\input{experiments}
%===============================================================================

%===============================================================================
\input{related_work}
%===============================================================================

%===============================================================================
\input{conclusion}
%===============================================================================
%\newpage
\bibliographystyle{splncs03}
\bibliography{biblio.bib}

\newpage
\input{appendix}
%\extendedonly{
%\input{appendix}
%}

\end{document}
