\section{Related Work}
%
The work of~\cite{franzle} presents a tight integration of SAT solving
with interval based arithmetic constraint solving to handle large constraint
systems. 
%
Silva et al.~\cite{sas12} present an abstract interpretation account of 
satisfiability algorithms derived from DPLL procedures.  
%
The work of~\cite{tacas12} is a very early instantiation of abstract 
CDCL~\cite{sas12} as an interval-based decision procedure for programs, 
but in a purely logical settings.  
%
A similar technique that lifts DPLL(T) to programs is Satisfiability Modulo 
Path Programs (SMPP)~\cite{SMPP}. SMPP enumerates program paths using a SAT 
formula, which are then verified using abstract interpretation.  
%
%The work of \cite{DBLP:conf/esop/MineBR16} proposes an algorithm inspired by 
%constraint solvers for inferring disjunctive invariants using intervals.
%
The lifting of CDCL to first-order theories is proposed
in~\cite{dpll,ndsmt,vmcai13}.
%
\Omit{ operates on a fixed first-order partial assignment lattice
  structure, where first-order variables are mapped to domain values,
  similar to constants lattice in program analysis.  } 
  More recently, ~\cite{cade17} proposed a calculus for extending theory of
  finite sets to theory of finite relations in CVC4 solver, which allows
  relational constraint solving in SMT.  
  \Omit{Unlike previous work that operates on a fixed first-order lattice,
  ACDLP can be instantiated with different abstract domains as well as product
  domains.  
  This involves model search and learning in abstract lattices.  A similar
  technique that lifts decisions, propagations and learning to theory variables
  is Model-Constructing Satisfiability Calculus (mcSAT)~\cite{vmcai13}.}
  
  ACDLP is not, however, similar to abstraction refinement. ACDLP works on a fixed
  abstraction. Also, transformer learning in ACDLP does not soundly over-approximate
  the existing program transformers. Hence, transformer learning in ACDLP is
  distinct from transformer refinement in classical CEGAR. 

\Omit { The
  abstract lattice in natural domain SMT does not have complementable
  meet irreducibles, and therefore does not support generalized clause
  learning~\cite{sas}.  On the other hand, most abstract lattice have
  complementation property, thus enabling ACDCL to perform generalized
  clause learning.  } 
%
%\cite{DBLP:journals/fmsd/BrainDGHK14} presents a similar decision and learning
%based approach for heap-manipulating programs, but uses a simplistic conflict analysis.
%
%\cite{DBLP:conf/vmcai/PelleauMTB13} describes a constraint solving algorithm
%using relational numerical domains, but performs model search only.
%
%
% is presented in . They
%use a relational domain, but focus on the model search, whereas the
%conflict analysis is simplistic and restarts the algorithm with the
%learnt clause instead of backtracking.
%}

